---
title: "Assignment 2: Coding Basics"
author: "Elsie Liu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Basics Day 1

```{r}
#1. Generate a sequence of numbers from 1 to 100, increasing by fours. Name this sequence as "seq_1_4_100".
seq_1_4_100 <- seq(1,100,4)

#2. Calculate the mean and median of the sequence.
mean(seq_1_4_100)
median(seq_1_4_100)

#3. determine whether the mean is greater than the median
mean(seq_1_4_100) > median(seq_1_4_100)
```

## Basics Day 2

```{r}
#5. Create a series of vectors, each with four components, consisting of 
#(a) names of students,
#(b) test scores out of a total 100 points
#(c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.
stdname <- c("Tom", "Todd", "May", "Merry")
scr <- c(48, 96, 96, 48)
ispass <- scr>50

#6. Label each vector with a comment on what type of vector it is.
class(stdname)
class(scr)
class(ispass)

#7. Combine each of the vectors into a data frame. Assign the data frame an informative name.
test.result <- data.frame(stdname,scr,ispass)

#8. Label the columns of your data frame with informative titles.
colnames(test.result) <- c("Name", "Score", "Pass or Not")
```

9. QUESTION: How is this data frame different from a matrix?

> **Answer**: The elements in a matrix are in the same format (e.g. numeric), while one data frame could hold varoius types of data.

```{r}
#10. Create a function with an if/else statement. Your function should determine whether a test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. Hint: Use `print`, not `return`. The name of your function should be informative.
is.pass.50 <- function(x) {
  for (i in 1:length(x)){
  if (x[i]>50) {
    print("TRUE")
  }
  else {
    print("FALUSE")
  }}
}

is.pass.50.ifelse <-function(x) {
  ifelse(x>50,"TURE","FALSE")
}

#11. Apply your function to the vector with test scores that you created in number 5. 
is.pass.50(scr)
is.pass.50.ifelse(scr)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked?

> **Answer**: Both of the function worked. "ifelse" is more convenient as it could handle the whole vector instead of only one element.
