---
title: "Assignment 09: Data Scraping"
author: "Elsie Liu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

# Total points:

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Set up 
1. Set up your session:

* Check your working directory
* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Set your ggplot theme

```{r, message = FALSE}
#1
getwd()
library(tidyverse)
library(lubridate)
library(rvest)
library(ggplot2)
library(plyr)
mytheme <- theme_classic() +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom",plot.title = element_text(hjust = 0.5))
theme_set(mytheme)
```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2020 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Change the date from 2021 to 2020 in the upper right corner.
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2019>
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website}
#2
DurhamURL <- read_html("https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2020")
```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PSWID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Max Daily Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to three separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values, with the first value being 36.0100.

```{r scrape.the.data}
#3
water.system.name <- DurhamURL %>% 
  html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>% 
  html_text()
pswid <- DurhamURL %>% 
  html_nodes("div+ table tr:nth-child(1) td:nth-child(5)") %>% 
  html_text()
ownership <- DurhamURL %>% 
  html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>% 
  html_text()
max.withdrawals.mgd <- DurhamURL %>% 
  html_nodes("th~ td+ td") %>% 
  html_text()
```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in order. You can overcome this by creating a month column in the same order the data are scraped: Jan, May, Sept, Feb, etc...

5. Plot the max daily withdrawals across the months for 2020

```{r create.a.dataframe.from.scraped.data}
#4
DWaterData <- data.frame("SystemName"=rep(water.system.name,12),"PSWID"=rep(pswid,12),"Ownership"=rep(ownership,12),"MGD"=max.withdrawals.mgd,"Month"=c(01,05,09,02,06,10,03,07,11,04,08,12),"Year"=rep(2020,12))
DWaterData$Date <- ym(with(DWaterData,paste(Year,Month,sep="-")))
#5
ggplot(data = DWaterData,aes(x = Date,y = MGD,group=1))+
  geom_line(color="lightblue")+
  geom_point()+
  ggtitle("Max daily withdrawals in Durham 2020")

```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site scraped**.

```{r construct.a.scraping.function}
#6.

waterscraping <- function(siteID,Year){
  try(if(class(siteID)!="character") stop("Need siteID as string"))
  if (Year==2020) {
  URL <- read_html(paste("https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=", siteID,sep = ""))
  } 
  else {
  URL <- read_html(paste("https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=", siteID, "&year=", Year,sep = ""))
  }
  water.system.name <- URL %>% 
    html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>% 
    html_text()
  pswid <- URL %>% 
    html_nodes("div+ table tr:nth-child(1) td:nth-child(5)") %>% 
    html_text()
  ownership <- URL %>% 
    html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>% 
    html_text()
  max.withdrawals.mgd <- URL %>% 
    html_nodes("th~ td+ td") %>% 
    html_text()
  WaterData <- data.frame("SystemName"=rep(water.system.name,12),"PSWID"=rep(pswid,12),"Ownership"=rep(ownership,12),"MGD"=max.withdrawals.mgd,"Month"=c(01,05,09,02,06,10,03,07,11,04,08,12),"Year"=rep(Year,12))
  WaterData$Date <- ym(with(WaterData,paste(Year,Month,sep="-")))
  return(WaterData)
}

```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data}
#7
Durham15<- waterscraping(siteID = '03-32-010',2015)
Durham15
```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares the Asheville to Durham's water withdrawals.

```{r }
#8
Asheville15<- waterscraping(siteID = '01-11-010',2015)
Asheville15
DurAsh15 <- full_join(Durham15,Asheville15)
DurAsh15 <- reshape(DurAsh15, idvar = "Month", v.names = "MGD", timevar = "SystemName", direction = "wide")

ggplot(data = DurAsh15)+
  geom_point(aes(x = Date,y = MGD.Durham,color="blue"))+
  geom_line(aes(x = Date,y = MGD.Durham,color="blue",group=1))+
  geom_point(aes(x = Date,y = MGD.Asheville,color="red"))+
  geom_line(aes(x = Date,y = MGD.Asheville,color="red",group=1))+
  labs(title = "Water withdrawals in Durham & Ahseville in 2015", x = "Month", y = "MGD", color = "County") +
  scale_color_manual(labels = c("Durham", "Asheville"),values = c("blue","red"))
```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2019.Add a smoothed line to the plot.

```{r}
#9
Ash10 <- waterscraping('01-11-010',2010)
Ash11 <- waterscraping('01-11-010',2011)
Ash12 <- waterscraping('01-11-010',2012)
Ash13 <- waterscraping('01-11-010',2013)
Ash14 <- waterscraping('01-11-010',2014)
Ash15 <- waterscraping('01-11-010',2015)
Ash16 <- waterscraping('01-11-010',2016)
Ash17 <- waterscraping('01-11-010',2017)
Ash18 <- waterscraping('01-11-010',2018)
Ash19 <- waterscraping('01-11-010',2019)
Ash1019 <- rbind(Ash10,Ash11,Ash12,Ash13,Ash14,Ash15,Ash16,Ash17,Ash18,Ash19)
Ash1019$MGD <- as.numeric(Ash1019$MGD)

par(mfrow=c(2,2))
ggplot(data = Ash1019,aes(x = Date, y = MGD))+
  geom_point()+
  geom_smooth(method = lm)+
  scale_x_date(date_labels = "%Y",date_breaks = "1 year")+
  labs(title = "Asheville's max daily withdrawal by months, 2010~2019")+
  theme(axis.text.x = element_text(size = 8))

ggplot(data = Ash1019)+
  geom_line(aes(x = Month, y = MGD, group=Year,color=as.factor(Year)))+
  scale_y_continuous(breaks = seq(19, 29, by = 1))+
  scale_x_continuous(breaks = seq(1, 12, by = 1))+
  labs(title = "Asheville's max daily withdrawal by months, 2010~2019")
```

>Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?

> **Answer**: Asheville has a increasing trend in water usage over years and also has a seasonal trend within each year.