---
title: "Assignment 6: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Elsie Liu"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

The completed exercise is due on Monday, February 28 at 7:00 pm.

## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r}
#1
getwd()
lter <- read.csv("./Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv")
lter$sampledate <- as.Date(lter$sampledate, format = "%m/%d/%y")
library(ggplot2)
library(dplyr)
library(corrplot)
library(agricolae)
#2
mytheme <- theme_classic(base_size = 10) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
theme_set(mytheme)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?   

3. State the null and alternative hypotheses for this question:   
**Answer**:   
$H_0$: the mean lake temperature recorded during July remains the same at all depth across all lakes;   
$H_a$: the mean lake temperature recorded during July changes with depth across all lakes.   


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4
lter.july <- 
  lter %>%
  filter(months(sampledate)=="July") %>%
  select(lakename, year4, daynum, depth, temperature_C) %>%
  na.omit()

#5
ggplot(data = lter.july,aes(x = depth, y = temperature_C))+
  geom_point(shape=1)+
  geom_smooth(method = lm) +
  ylim(0,35)+
  ylab("Temperature")+
  xlab("Depth")

```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?   

**Answer**: According to the figure, the temperature recorded in July decreases with the depth.   


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7
lm.depth <- lm(data = lter.july, formula = temperature_C~depth)
summary(lm.depth)
```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth.     

**Answer**: The linear relationship between depth and temperature is significant (P<0.05). About 73.87% of the variance is explained by depth (DF=9726). The temperature is predicted to decrease 1.94621 degrees for every 1 m growth in depth.


---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 

9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9
ltercor<-cor(lter.july[,-1])
corrplot.mixed(ltercor,upper = "ellipse")
lm.AIC <- lm(data = lter.july, formula = temperature_C ~ year4 +daynum+depth)
step(lm.AIC)
#10
#year+daynum+depth is the best set
summary(lm.AIC)

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?   

**Answer**: The AIC method suggests that the model with all the three variables (year4, daynum, depth) best predicts temperature. About 74.11% of the variance is explained by the model. This multiple regression model slightly increased the proportion of variance explained.

---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
lter.anova <- aov(data=lter.july, temperature_C ~ lakename)
summary(lter.anova)
lter.anova2 <- lm(data=lter.july, temperature_C ~ lakename)
summary(lter.anova2)
```

13. Is there a significant difference in mean temperature among the lakes? Report your findings.    

**Answer**: Yes, there is a significant difference in mean temperature among lakes. ANOVA tests showed that we should reject null hypothesis and the mean temperature for different lakes are not equal.

14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.
ggplot(data = lter.july,aes(x = depth,y = temperature_C,color=lakename))+
  geom_point(alpha=0.5)+
  geom_smooth(method = lm)+
  ylim(0,35)+
  ylab("Temperature")+
  xlab("Depth")+
  guides(fill = guide_legend(title = "Lake name"))+
  labs(title = "Lake temperature by depth")
```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15
TukeyHSD(lter.anova)
par(mar=c(2,9,2,2))
plot(TukeyHSD(lter.anova),cex.axis=0.5,las=1)
lterHSD <- HSD.test(lter.anova, "lakename", group = T)
lterHSD
```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?    
**Answer**: Paul Lake & Ward Lake statistically has the same mean temerature as Peter Lake. No lake has statistically distinct temperature from other lakes.    
    
17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures?             
**Answer**: The two sample T-test.
